---
title: "Homework 4"
author: "Brian Bruxvoort"
date: "2023-11-02"
output: pdf_document
---

# Page 189, Question 12, Part a
```{r}
library(MASS)
cereal <- read.csv("cereal_dillons.csv")
stand01 <- function(x) {(x - min(x)) / (max(x) - min(x))}
cereal2 <- data.frame(Shelf = cereal$Shelf, sugar =
  stand01(x = cereal$sugar_g / cereal$size_g), fat =
  stand01(x = cereal$fat_g / cereal$size_g), sodium =
  stand01(x = cereal$sodium_mg / cereal$size_g))
```


# Part b
```{r}
boxplot(formula = sugar ~ Shelf, data = cereal2, ylab = "Sugar", xlab = "Shelf", pars = list(outpch = NA))
stripchart(x = cereal2$sugar ~ cereal2$Shelf, lwd = 2 , col = "red", method = "jitter", vertical = TRUE, pch = 1, add = TRUE)
```

```{r}
cereal3<-data.frame(cereal2[,2:4])

# Colors by condition:
cereal.colors<-ifelse(test = cereal2$Shelf==1, yes = "purple", 
                no = ifelse(test = cereal2$Shelf==2, yes = "blue", 
                no = ifelse(test = cereal2$Shelf==3, yes = "green", no= "red")))
# Line type by condition:
cereal.lty<-ifelse(test = cereal2$Shelf==1, yes = "solid", 
                no = ifelse(test = cereal2$Shelf==2, yes = "twodash", 
                no = ifelse(test = cereal2$Shelf==3, yes = "solid", no = "twodash")))

parcoord(x = cereal3, col = cereal.colors, lty = cereal.lty)  # Plot
legend(x = 3.35, y = 1.05, legend = c("Shelf 1", "Shelf 2", "Shelf 3", "Shelf 4"), lty = c("solid", "twodash", "solid", "twodash"),
  col=c("purple", "blue", "green","red"), cex=0.8, bty="n")
```

High sugar seems to be prevalent in shelf 2 and high sodium seems to be prevalent in shelf 1.

# Part c
If we believed there was a natural ordering of the shelves and the cereal on them, then we would consider ordinality. However, for this situation the shelves are not naturally ordered, so we would not consider ordinality for this example.

# Part d
```{r}
library(package = nnet)
mod.fit <- multinom(formula = as.factor(Shelf) ~ sugar + fat + sodium, data = cereal2)
summary(mod.fit)
```

```{r}
library(package = car)
Anova(mod = mod.fit)
```



# Part e
```{r}
mod.fit2 <- multinom(formula = Shelf ~ sugar + fat + sodium + sugar:fat + sugar:sodium + fat:sodium + sugar:fat:sodium, data = cereal2)
summary(mod.fit2)
```

```{r}
Anova(mod.fit2)
```




# Part f
```{r}
stand.new <- function(meas, serv_size, comparison) {(meas/serv_size - min(comparison)) / (max(comparison) - min(comparison))}

newdata <- data.frame(sugar = stand.new(12, 28, cereal$sugar_g/cereal$size_g), 
                      fat = stand.new(0.5, 28, cereal$fat_g/cereal$size_g),
                      sodium = stand.new(130, 28, cereal$sodium_mg/cereal$size_g))

round(predict(object = mod.fit, newdata = newdata, type = "probs", se.fit = TRUE),7)
```


# Part g
```{r}
beta.hat<-coefficients(mod.fit)
beta.hat
mean_fat <- mean(cereal2$fat)
mean_sodium <- mean(cereal2$sodium)


# Create plotting area first to make sure get the whole region with respect to x-axis
curve(expr = 1/(1 + exp(beta.hat[1,1] + beta.hat[1,2]*x) + exp(beta.hat[2,1] + beta.hat[2,2]*x)), ylab = expression(hat(pi)), xlab = "sugar",
  xlim = c(min(cereal2$sugar), max(cereal2$sugar)), ylim = c(0, 1), col = "black", lty = "solid", lwd = 2, n = 1000, type = "n",
  panel.first = grid(col = "gray", lty = "dotted"))

## Plot each pi_j
# Shelf1
curve(expr = 1/(1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)),
  col = "black", lty = "solid", lwd = 2, n = 1000, add = TRUE,
  xlim = c(min(cereal2$sugar[cereal2$Shelf == 1]), max(cereal2$sugar[cereal2$Shelf == 1])))

curve(expr = 1/(1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)),
  col = "black", lty = "dotdash", lwd = 2, n = 1000, add = TRUE,
  xlim = c(0,1))

# Shelf2
curve(expr = exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)/
        (1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)), 
      col = "green", lty = "solid", lwd = 2, n = 1000, add = TRUE, 
      xlim = c(min(cereal2$sugar[cereal2$Shelf == 2]), max(cereal2$sugar[cereal2$Shelf == 2])))

curve(expr = exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)/
        (1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)), 
      col = "green", lty = "dotdash", lwd = 2, n = 1000, add = TRUE, 
      xlim = c(0,1))

# Shelf3
curve(expr = exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)/
        (1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)),
  col = "red", lty = "solid", lwd = 2, n = 1000, add = TRUE,
  xlim = c(min(cereal2$sugar[cereal2$Shelf == 3]), max(cereal2$sugar[cereal2$Shelf == 3])))

curve(expr = exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)/
        (1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)),
  col = "red", lty = "dotdash", lwd = 2, n = 1000, add = TRUE,
  xlim = c(0,1))

# Shelf4
curve(expr = exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)/
        (1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)),
  col = "blue", lty = "solid", lwd = 2, n = 1000, add = TRUE,
  xlim = c(min(cereal2$sugar[cereal2$Shelf == 4]), max(cereal2$sugar[cereal2$Shelf == 4])))   

curve(expr = exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)/
        (1 + exp(beta.hat[1,1] + beta.hat[1,2]*x + beta.hat[1,3]*mean_fat + beta.hat[1,4]*mean_sodium)
                + exp(beta.hat[2,1] + beta.hat[2,2]*x + beta.hat[2,3]*mean_fat + beta.hat[2,4]*mean_sodium)
                + exp(beta.hat[3,1] + beta.hat[3,2]*x + beta.hat[3,3]*mean_fat + beta.hat[3,4]*mean_sodium)),
  col = "blue", lty = "dotdash", lwd = 2, n = 1000, add = TRUE,
  xlim = c(0,1))




legend(x = 0.5, y = 1.0, legend=c("Shelf 1", "Shelf 2", "Shelf 3", "Shelf 4"), 
       lty=c("solid","solid", "solid", "solid"),
  col=c("black","green","red", "blue"), bty="n", lwd = c(2,2,2), seg.len = 4)
```



# Part h
```{r}
sd.cereal<-apply(X=cereal2[, -c(2)], MARGIN = 2, FUN = sd)
c.value<-c(sd.cereal)[2:4]
# Estimated standard deviations for each explanatory variable
round(c.value,2)

conf.beta <- confint(object = mod.fit, level = 0.95)
ci.OR <- exp(c.value*conf.beta[2:4,1:2,])

#coefficients(mod1)
beta.hat2<-coefficients(mod.fit)[1,2:4]
beta.hat3<-coefficients(mod.fit)[2,2:4]
beta.hat4<-coefficients(mod.fit)[3,2:4]

# OR for j = 2 (Shelf 2 vs Shelf 1)
print("OR for j = 2 vs j = 1")
mid = exp(c.value*beta.hat2)
round(mid,4)
round(1/mid,4)
round(data.frame(low = ci.OR[,1,1], mid = mid, up = ci.OR[,2,1]),4)

# OR for j = 3 (Shelf 3 vs Shelf 1)
print("OR for j = 3 vs j = 1")
mid = exp(c.value*beta.hat3)
round(mid,4)
round(1/mid,4)
round(data.frame(low = ci.OR[,1,2], mid = mid, up = ci.OR[,2,2]),4)
# Significant variables
round(data.frame(low = 1/ci.OR[c(1,3),2,2], mid = 1/mid[c(1,3)], up = 1/ci.OR[c(1,3),1,2]),4)

# OR for j = 4 (Shelf 4 vs Shelf 1)
print("OR for j = 3 vs j = 1")
mid = exp(c.value*beta.hat4)
round(mid,4)
round(1/mid,4)
round(data.frame(low = ci.OR[,1,3], mid = mid, up = ci.OR[,2,3]),4)
# Significant variables
round(data.frame(low = 1/ci.OR[c(1,3),2,3], mid = 1/mid[c(1,3)], up = 1/ci.OR[c(1,3),1,3]),4)

cereal2$new_shelf <- relevel(as.factor(cereal2$Shelf),"2")
mod.fit <- multinom(new_shelf ~ sugar + fat + sodium, data = cereal2)
conf.beta.new <- confint(object = mod.fit, level = 0.95)
ci.OR.new <- exp(c.value*conf.beta.new[2:4,1:2,])
beta.hat3.new<-coefficients(mod.fit)[2,2:4]
beta.hat4.new<-coefficients(mod.fit)[3,2:4]

# OR for j = 3 (Shelf 3 vs Shelf 2)
print("OR for j = 3 vs j = 2")
mid = exp(c.value*beta.hat3.new)
round(mid,4)
round(1/mid,4)
round(data.frame(low = ci.OR.new[,1,2], mid = mid, up = ci.OR.new[,2,2]),4)
# Significant variables
round(data.frame(low = 1/ci.OR.new[1,2,2], mid = 1/mid[1], up = 1/ci.OR.new[1,1,2]),4)

# OR for j = 4 (Shelf 4 vs Shelf 2)
print("OR for j = 4 vs j = 2")
mid = exp(c.value*beta.hat4.new)
round(mid,4)
round(1/mid,4)
round(data.frame(low = ci.OR.new[,1,3], mid = mid, up = ci.OR.new[,2,3]),2)
# Significant variables
round(data.frame(low = 1/ci.OR.new[1,2,3], mid = 1/mid[1], up = 1/ci.OR.new[1,1,3]),4)

cereal2$new_shelf <- relevel(as.factor(cereal2$Shelf),"3")
mod.fit <- multinom(new_shelf ~ sugar + fat + sodium, data = cereal2)
conf.beta.new <- confint(object = mod.fit, level = 0.95)
ci.OR.new <- exp(c.value*conf.beta.new[2:4,1:2,])
beta.hat4.new<-coefficients(mod.fit)[3,2:4]

# OR for j = 4 (Shelf 3 vs Shelf 3)
print("OR for j = 4 vs j = 3")
mid = exp(c.value*beta.hat4.new)
round(mid,4)
round(1/mid,4)
round(data.frame(low = ci.OR.new[,1,3], mid = mid, up = ci.OR.new[,2,3]),4)
```



